Paper

### 2024-12-13 paper/index.qmd
deterministic version

$$
\frac{dH_S}{dt} = A_S - QH_S(t)H_I(t) - DH_S(t)
$$

$$
\frac{dH_I}{dt} = A_I + QH_S(t)H_I(t) - DH_I(t)
$$

@@

>>>>

$$
\frac{dH_S}{dt} = A_S - Q\frac{H_S(t)}{H_I(t)} - DH_S(t)
$$

$$
\frac{dH_I}{dt} = A_I + Q\frac{H_S(t)}{H_I(t)} - DH_I(t)
$$

@@

```{r}
single_outbreak_sim_mod <- single_outbreak_sim %>% 
  mutate(prop_infected = `current-colonised` / `current-inpatients`,
         current_not_infected = `current-inpatients` - `current-colonised`)
single_outbreak_sim_mod %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  ggplot(aes(x = `[step]`, y = value, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment")

# single_outbreak_sum_stat <- single_outbreak_sim_mod %>% 
#   group_by(`[step]`) %>%
#   summarise(m = mean(prop_infected))
```

```{r}
# Parameters and time-dependent functions
parameters <- list(
  p = 1 - unique(single_outbreak_sim$`community-colonisation-rate`),      # Proportion of non-infected admissions
  A = 50        # Total admission/discharge rate
)
# 
# baselinerate <- seq(from=0.01, to=0.99, length.out = 5)
# outbreakrate <- seq(0.01, 0.99, length.out = 5)
# rate_grid <- expand.grid(baselinerate, outbreakrate)

# Time-varying beta function (example: sinusoidal variation over time)
beta_function <- function(t, baseline_rate, outbreak_rate) {
  if (t > unique(single_outbreak_sim$`outbreak-start`) &
      t < unique(single_outbreak_sim$`outbreak-end`)) {
    return (outbreak_rate)
  }
  return (baseline_rate)
}

# Initial conditions
initial_state <- c(N = 600, I = 71)  # Initial populations of non-infected and infected patients
time <- seq(0, run_to, by = 1)         # Time steps
dt <- 1                             # Time step size
# 
# plan(multisession)
# furrr::future_map2(rate_grid$Var1, rate_grid$Var2,
#         \(x, y) {
#           print("1")
#           N <- numeric(length(time))
#           I <- numeric(length(time))
#           N[1] <- initial_state["N"]
#           I[1] <- initial_state["I"]
#           
#           # Euler method for numerical integration with time-varying parameters
#           for (t in 1:(length(time) - 1)) {
#             # Current beta value
#             beta_t <- beta_function(time[t], x, y)
#             
#             # Total population at current time
#             total <- N[t] + I[t]
#             
#             # Proportions of non-infected and infected
#             prop_N <- ifelse(total > 0, N[t] / total, 0)
#             prop_I <- ifelse(total > 0, I[t] / total, 0)
#             
#             # Calculate derivatives
#             dN <- parameters$p * parameters$A - parameters$A * prop_N - beta_t * N[t] * prop_I
#             dI <- (1 - parameters$p) * parameters$A - parameters$A * prop_I + beta_t * N[t] * prop_I
#             
#             # Update populations using Euler's method
#             N[t + 1] <- N[t] + dN * dt
#             I[t + 1] <- I[t] + dI * dt
#           }
#           I[31:length(I)] / ( N[31:length(N)] + I[31:length(I)] )
#         })

baselinerate <- 0.02
outbreakrate <- 0.07

# Initialize vectors to store results
N <- numeric(length(time))
I <- numeric(length(time))
N[1] <- initial_state["N"]
I[1] <- initial_state["I"]

# Euler method for numerical integration with time-varying parameters
for (t in 1:(length(time) - 1)) {
  # Current beta value
  beta_t <- beta_function(time[t], baselinerate, outbreakrate)

  # Total population at current time
  total <- N[t] + I[t]

  # Proportions of non-infected and infected
  prop_N <- ifelse(total > 0, N[t] / total, 0)
  prop_I <- ifelse(total > 0, I[t] / total, 0)

  # Calculate derivatives
  dN <- parameters$p * parameters$A - parameters$A * prop_N - beta_t * N[t] * prop_I
  dI <- (1 - parameters$p) * parameters$A - parameters$A * prop_I + beta_t * N[t] * prop_I

  # Update populations using Euler's method
  N[t + 1] <- N[t] + dN * dt
  I[t + 1] <- I[t] + dI * dt
}

# Combine results into a data frame for plotting
output_df <- data.frame(
  time = time,
  N = N,
  I = I
)

output_df <- filter(output_df, time > 31)

# Plot the results
library(ggplot2)
output_long <- reshape2::melt(output_df, id = "time")
ggplot(output_long, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1) +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("Non-Infected (N)", "Infected (I)")) +
  theme_minimal()

single_outbreak_sim_mod %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  ggplot(aes(x = `[step]`, y = value, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  geom_line(data = output_df, aes(x = time, y = N), color = "blue") + 
  geom_line(data = output_df, aes(x = time, y = I), color = "red")
```
